Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-11-28 patchy-review-responses.tex

  (i) There needs to be some sort of numerical verification (e.g. with
  individual-based simulations) to test the main conclusions, check their
  robustness, and identify the parameter ranges for which they hold.
\end{quote}

We have done a large number of simulations, and compared the results to the theory;
the main comparison occurs in \ref{fig:sim_times} and figures~\ref{fig:sim_probs};
we have also added in supplementary section \ref{apx:parameter_tables} 
visualizations of the simulations at various parameter values
that we think will help some readers in verifying the assumptions.

\begin{quote}
  (ii) The presentation of the main result needs to be improved. In particular,
  all reviewers feel that the main result can be derived in a simpler and more
  intuitive way. I agree with them that at present the manuscript is overly
  technical and would benefit strongly from a more intuitive derivation of the
  results and discussion of the assumptions under which they hold. The reviewers
  provide detailed suggestions on how this can be achieved.
\end{quote}

We sympathize entirely: we felt the same way ourselves,
and went around in similar circles for quite some time
before settling on the argument given in the first draft.
As suggested by the third reviewer, 
we have moved the most difficult part of the argument to the Materials \& Methods,
replacing it in the Results with a more \textit{ad hoc} argument
along the lines of that suggested by the reviewers.
(We do think the more formal argument is more solid, however,
and therefore necessary to keep; 
as discussed below, various seemingly reasonable arguments
give different answers for subtle reasons.)

@@

Other changes of note are:
\begin{itemize}

  \item Several reviewers pointed out that in approximating $\E[1-(1-p_e)^K] \approx \E[K] p_e$
    we were neglecting the case that $s_m \ll s_p$.
    This omission was pure laziness on our part,
    which we have rectified; see the new equation~\eqref{eqn:twocase_migrate}.

  \item We have also improved our discussion of the limits to the theory,
    mostly in the (short) section \nameref{ss:assumptions}.
    The main point of note was that, indeed as Reviewer 3 pointed out,
    we were running tests in regions where $s_m$ was too small for the theory to apply,
    resulting in $\migrate$ that was not monotonic with $s_m$.
    Happily, the natural restriction of $R \sqrt{s_m}/\sigma > 1$ coincides with the condition
    for $\migrate$ to decrease with $s_m$;
    we don't have good intuition for why the formula should increase in the region where it is inapplicable,
    but this seems beside the point.

@@

    The latter is the case discussed in the manuscript. The other is important too
    and should be discussed. 
}

\reply
As discussed above, we have incorporated this good point; see around equation~\eqref{eqn:twocase_migrate}.

%%%%%%%%
\point{}{ 
    I don't seem to be able to identify the equivalent of
    $\ell$ in the author's formulae, but I believe it should be there. After all, 
    $\rho q(x)$ is a density and needs to be multiplied by a length (or area in d=2) to
    give a meaningful rate.
}

\reply
    The correspondence in \eqref{eqn:ratio_K} boils down to \eqref{eqn:boils_downto};
    but rearranging this to $u(x,S) \approx h(x,S) \E[K] g(A)$,
    this says that the mean occupation density in $S$ due to a family beginning at $x$
    is the mean number of lineages hitting $S$ ($h(x,S) \E[K]$)
    multiplied by the mean occupation density of a single lineage starting on the edge of the patch.
    So, it seems that the missing length (or area) enters through $g(A)$,
    which in our case is $1/(2s_m)$.
    We agree, it is not obvious how the units work out here.

@@

%%%%%%%%
\point{}{ 
    To assess the accuracy of the various approximations made, the process should
    be simulated and the measured rates explicitly compared to analytical
    predictions.
}

\reply
  We've done this; see above.

%%%%%%%%
\point{long range migration}{All results depend exponentially on the distance, which
      is a consequence of the model only allowing for short range dispersal. Rare
    long range dispersal will alter all these rates dramatically, see for example
    recent work by Hallatschek and Fisher.
}

\reply
This was already discussed in the discussion, but we've added a note about it earlier on, in particular adding a citation to the nice paper of Hallatschek \& Fisher.
\llname{rr:longtailed}

@@

%%%%%%%%
\point{Fig 4:}{ Why not use natural units for the axis of these plots,
  $R\sqrt(s)/\sigma$, $\rho \mu A/s$.
  }

\reply
Good idea, but we've decided to cut this figure.

%%%%%%%%
\point{in the discussion on page 10 "probability of parallel adaptation.."}{
    It should be noted that many of the scenarios at high mutation influx are quite
    hypothetical. While the colonization of migrants might dominate over local
    adaptation, this argument assumes that the migrants are at equilibrium
    frequency. The transient to reach this equilibrium might be much longer than
    the time required for de novo local adaptation.
}

\reply
We don't know of even a poor approximation to the time to reach migration-selection equilibrium,
unfortunately.
However, we have pointed out in \nameref{ss:assumptions} that we're missing this point,
and discuss there how the time lag to reach this equilbrium contributes to discrepancies between theory and simulations.

@@

%%%%%%%%%%%%%%  Reviewer 2
\reviewersection

\begin{quote}
Briefly, I think that the derivation of the main results (12), (21), and (25) can
be made much shorter and easier to understand (for both empiricists and the-
orists), and that there need to be simulations that quantitatively confirm these
results. Overly-detailed comments are below.
\end{quote}

\point{}{
    I think that the main result (12) can be derived much more simply, and
    also extended. By writing that the probability that a family that arrives
    at the patch is successful is approximately $ 2s_p \E [K] /\xi^2$, the authors have
    implicitly assumed that this quantity is $<\ll 1$. But this implies that the
    fluctuations in the number of B individuals arriving at the patch can be
    ignored, since even a large fluctuation does not saturate the probability
    of success. This is valid for $s_m \gg s_p$. Given that the fluctuations can be
    ignored, (12) can be derived immediately from $q(y)$ -- simply apply Fick’s
    Law to $q(y)\rho$ to get that the flux of new B individuals is $\sigma^2 |q'(y)|\rho/2$, and
    then multiply by the success probability $2s_p/\xi^2$. (Actually, the result is
    slightly different from (12) -- see X below.)
}

\reply
This argument implicitly treats each migrant individual as independent,
ignoring the correlations of arrival times induced by families.
This is the reason for explicitly defining families to be those who share an ancestor living outside of $r_0$
more recently than one living inside the patch;
many of the particles that Fick's law would pick up leaving $r_0$ would be already counted as part of a family.
However, we think our simplified argument in \nameref{ss:heuristics} gets the same spirit.

@@

\point{}{
    However, I think we should also care about the opposite limit, $s_m \ll s p$. In
    this case, the fluctuations do matter -- arrivals typically occur in clumps
    (families) large enough so that success is almost certain conditioned on
    arrival. Since on average each family consists of $\E[K]$ individuals, the
    family flux is just the individual flux divided by $\E[K]$. $\E[K] = \xi^2 /(2s_m)$,
    so the result is the same as above, except with $s_p$ replaced by an additional
    factor of $s_m$. (This could be made more formal, but I don’t think that’s
    what the authors are going for.) In the intermediate case $s_m \approx s_p$ which
    might be the most relevant for the mice, the two limiting regimes cross
    over smoothly. So for the first limit ($s_m \gg s_p$), all you need is $q$, and
    for the second, you only need to add basic facts about branching random
    walks (ie, limiting number distribution and localization at long times).
}

\reply
Good point about $s_m \ll s_p$ (also made by reviewer \#1);
we've modified the key equation \eqref{eqn:migrate}
and surrounding discussion along these lines.

@@

%%%%%%%%
\point{}{ 
    I think that the authors need to show simulation results confirming the
    analytic approximations. Spatial simulations are a pain, but it’s important
    to have a (somewhat) independent check on the math and assumptions. At
    a minimum, I think that there needs to be a figure showing that they have
    the right expression for the rate of successful migrants (12) and migration
    sojourn times (17) over a range of parameter values. If it’s not too much
    work, I think it would be really nice to have a figure that doubles as a kind
    of graphical summary, showing a scattering of patches in two dimensions,
    with one starting as the homeland of the B allele and the rest colored
    according to their probability in simulations of being colonized by that
    original allele vs new mutations, and a curve marking the region predicted
    as being colonized by the analytical approximations, or something like
    that. But I understand if the authors don’t want to spend their time on
    that.
}

\reply
As discussed above, we've done this.

%%%%%%%%
\point{}{ 
    It would be nice to have a table of symbol definitions.
}

\reply
We've tried to say in words what each symbol is when we use it as much as possible,
so we hope this isn't still necessary.

@@

\reply
Good point, but $A$ is what shows up naturally in $\mutrate$, so we've left it as-is.

%%%%%%%%
\point{}{ 
     It would be nice to have a clear, concise summary of the quantitative
     assumptions needed for the results (all the stuff like $a \gg \sigma/\sqrt{2s_m}$ or
     whatever).
 }

\reply
See text added in \nameref{ss:assumptions}.

%%%%%%%%
\point{"Length of the hitchhiking haplotype"}{
    I would focus more on the asymptotic expressions for $R \gg w  \gg \sigma/ \sqrt{2s_m}$, since that’s the regime
     where the analysis makes the most sense. Then the expressions become
     simpler and you can treat $d = 2$ as well. If you want, you could even
     lead with the intuitive explanation at the end of the section, and save the
     Bessel functions for an appendix.
}

@@

 \reply

%%%%%%%%
\point{}{ 
     I think (24) follows directly from Barton, Heredity 1979. (If you look at
     the unnumbered equation above Fig.\ 1 in that paper and plug in $x \sim a$,
     you find the time scale $r_{eff}$.) I think that the displayed equation should just
     be $r_{eff} \approx r\sigma/\sqrt{s}/a$. (I feel like the extra correction terms and geometric
      factor go against the spirit of the paper, and in any case are probably
      small compared to the inaccuracies introduced just by applying the model
      to any real system.) I think you can cut essentially everything before
      the simple explanation just above (24); if people want the mathematical
      details, they can read Barton. If I am misunderstanding things and there
    is a new result here, rather than just a good new exposition of an existing
    result, there should be a figure showing simulations confirming it.
}

\reply
This is a very interesting point,
and the basic tool of a diffusion approximation seems to be the same in both.
However, something is not matching up,
because \citet{barton1979geneflow} finds that the length of the region of differentiated genome 
decays with $1/\sqrt{T}$,
whereas with our approximation the length of the differentiated haplotype decays as $1/T$.
The difference may be that Barton considers LD and we consider a haplotype unbroken by recombination.
On the balance, and considering the other reviewer's wise comments,
we've cut most of this from \nameref{ss:patch_haplotype}.

### 2014-11-28 patchy-selection-paper.tex
.

\begin{figure}[ht!]
  \begin{center}
    \includegraphics{sim-occupation-freqs}
  \end{center}
  \caption{
  \textbf{(A)} 
  Mean occupation frequencies of individual demes (grey dots)
  and averaged over demes at similar distances (black circles), on
  a one-dimensional grid of 201 demes, and
  a two-dimensional grid of $101\times 101$ demes.
  Superimposed in color is the decay predicted by \eqref{eqn:eqfreq},
  with $C$ chosen arbitrarily to give a reasonable fit.
  \plr{what is $C$?}
  (The disagreement at long distances is due to boundary effects near the edge of the range.)
  \textbf{(B)} 
  Fluctuations: colored lines are snapshots of allele frequencies at evenly spaced times
  from a one-dimensional grid of 151 demes;
  the solid line shows the cumulative mean occupation frequency across all generations.
  For both plots, the simulation was run for 1000 generations;
  the focal allele is beneficial in the central 10 demes, with $s_p=0.1$,
  and deleterious elsewhere, with $s_m=-0.02$; 
  demes had 1000 individuals in each;
  dispersal was to nearby demes with a mean dispersal distance of $\sigma \approx 1.66$  times the deme spacing.
   }   \label{fig:sim_occupation_freqs}
\end{figure}

@@

To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
This is a collection of trees, each rooted at an allele living outside the patch whose parent was born inside the patch.
Next, fix some intermediate distance $r_0$ from the established patch,
and erase from the genealogy every allele that has no ancestors living further away than $r_0$ to the patch.
This fragments the original set of trees into a collection of smaller trees that relate to each other all $B$ alleles living outside of $r_0$ at any time,
and some living inside of $r_0$;
we refer to these trees as ``families''.
If $r_0$ is large enough that there are not too many families
and interactions between family members are weak,
then these families can be approximated 
by a collection of independent spatial branching processes
whose members are ignored if they enter the original patch.
(This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
In terms of these families, the equilibrium frequency above is then
\begin{align}
    \label{eqn:gestalt_q}
    \begin{split}
        q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
\end{split}
\end{align}
and the quantity we wish to compute,
the effective rate at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
is
\begin{align}
    \label{eqn:gestalt_migrate}
    \begin{split}
        \migrate(x) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $x$ ) }
    \end{split}
\end{align}
The quantities on the right-hand side depend implicitly on $r_0$,
but their product does not.
All terms except the ``outflux of families'' are calculations with spatial branching processes;
and we can solve for this outflux using the established form for $q(x)$.
% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.

@@

A quick review of the parameter regime we are studying here:
First, we assume that 
\textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
Next, in using equation \eqref{eqn:eqfreq}, we assume that 
\textbf{(b)} the inter-patch distance is at least the characteristic length
($R > \sigma/\sqrt{s_m}$),
and that
\textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
we cannot approximate its dynamics via a branching process.
We also neglect the time for migration-selection equilibrium to be reached,
which has the effect of over-estimating $\migrate$,
but don't know a good way to estimate this contribution.
As discussed above, we also assume that migration to a new patch takes place over a number of generations;
if there are sufficient rare, long-distance migration events
that would move between patches in a single hop,
this would require a different analysis.

@@

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      \includegraphics{times-predicted-observed}
  \end{center}
  \caption{
      \textbf{(left)} 
      Time until adaptation by mutation in an isolated patch for different parameter combinations,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
      \textbf{(right)} 
      Time until adaptation by migration from an already occupied patch to a new patch,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
      For each,
      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:migration_params} and \ref{stab:mutation_params};
      shown are the median (points) and interquartile range (lines)
      of the time until the patch had at least 100 adapted individuals,
      for each unique parameter combination.
  }   \label{fig:sim_times}
\end{figure}

@@

@@

First we briefly describe the simulations we used for illustration and validation
(the \texttt{R} code used is provided as a supplement).
We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
locally beneficial alleles were more likely to reproduce.
Each then produced a random, Poisson distributed number of offspring that each, independently, 
either remained in the same location with probabilty $0.8$ or
migrated a random number of steps in a uniformly chosen cardinal direction;
the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
\linelabel{rr:sim_details}
Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
(Although we described the simulation in terms of individuals,
we kept track only of total numbers in an equivalent way.)

@@

Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
then this is equal to
\begin{align}
    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
\end{align}
Therefore, we have that
\begin{align}
    h(x,S) \E[K] 
    \approx \P^x[ \tau_S < \tau_\dagger ] 
    = u(x,S) / g(A) .
\end{align}

>>>>

\begin{figure}[ht!]
  \begin{center}
    \includegraphics{sim-occupation-freqs}
  \end{center}
  \caption{
  \textbf{(A)} 
  Mean occupation frequencies of individual demes (grey dots)
  and averaged over demes at similar distances (black circles), on
  a one-dimensional grid of 201 demes, and
  a two-dimensional grid of $101\times 101$ demes.
  Superimposed in color is the decay predicted by \eqref{eqn:eqfreq},
  with $C$ chosen arbitrarily to give a reasonable fit: 0.85 and 0.11 respectively.
  (The disagreement at long distances is due to boundary effects near the edge of the range.)
  \textbf{(B)} 
  Fluctuations: colored lines are snapshots of allele frequencies at evenly spaced times
  from a one-dimensional grid of 151 demes;
  the solid line shows the cumulative mean occupation frequency across all generations.
  For both plots, the simulation was run for 1000 generations;
  the focal allele is beneficial in the central 10 demes, with $s_p=0.1$,
  and deleterious elsewhere, with $s_m=-0.02$; 
  demes had 1000 individuals in each;
  dispersal was to nearby demes with mean dispersal distances of 
  $\sigma = 0.95$ and 0.74 times the deme spacing, respectively.
   }   \label{fig:sim_occupation_freqs}
\end{figure}

@@

To make the argument, we decompose the genealogy of $B$ alleles into families in the following way.
First, consider the genealogy of all $B$ alleles that were alive at any time outside of the original patch.  
This is a collection of trees, each rooted at an allele living outside the patch whose parent was born inside the patch.
Next, fix some intermediate distance $r_0$ from the established patch,
and erase from the genealogy every allele that has no ancestors living further away than $r_0$ to the patch.
This fragments the original set of trees into a collection of smaller trees that relate to each other all $B$ alleles living outside of $r_0$ at any time,
and some living inside of $r_0$;
we refer to these trees as ``families''.
If $r_0$ is large enough that there are not too many families
and interactions between family members are weak,
then these families can be approximated 
by a collection of independent spatial branching processes
whose members are ignored if they enter the original patch.
(This can be made formal in the limit of large population density, also taking $r_0$ large enough that the chance of reaching the original patch is small.)
This decomposition is illustrated in figure~\ref{fig:branching_decomp}.
In terms of these families, the equilibrium frequency above is then
\begin{align}
    \label{eqn:gestalt_q}
    \begin{split}
        \rho q(x) &= \text{ ( outflux of families ) } \times \text{ ( mean occupation density of a family at $x$ ) } ,
\end{split}
\end{align}
and the quantity we wish to compute,
the effective rate at which \emph{families of} migrant $B$ alleles establish in a new patch at displacement $x$ from the original,
is
\begin{align}
    \label{eqn:gestalt_migrate}
    \begin{split}
        \migrate(x) &= \text{ ( outflux of families ) } \times \text{ ( probability a family establishes in patch at $x$ ) }
    \end{split}
\end{align}
The quantities on the right-hand side depend implicitly on $r_0$,
but their product does not.
All terms except the ``outflux of families'' are calculations with spatial branching processes;
and we can solve for this outflux using the established form for $q(x)$.
% Note that $\migrate$ reflects the mean number of establishing migrant families per unit time, rather than individuals
% -- a naive calculation using $q(x)$ that assumes migrants are independent would overestimate the effective migration rate.

@@

Now that we have expressions for the mean rates of adaptation by new mutation, \eqref{eqn:mutrate},
and by migration from an already colonized patch, \eqref{eqn:migrate},
it seems helpful to step back and review the assumptions we have made and the parameter regime
we expect the results to apply to.
For the mutation rate,
our results will hold best if $s_m \approx s_p$,
if the patch diameter is large relative to $\sigma/\sqrt{\min(s_m,s_p)}$,
the patch is not too far from circular.
If $s_m \ll s_p$ then \eqref{eqn:mutrate} will underestimate the rate,
and conversely if $s_m \gg s_p$.
As for the migration rate,
we assume that 
\textbf{(a)} each patch is large enough to support a stable population of $B$ alleles
($A^{1/d} > (\sigma/\sqrt{2s_p}) \tan^{-1}(\sqrt{s_m/s_p})$).
Next, in using equation \eqref{eqn:eqfreq}, we assume that 
\textbf{(b)} the inter-patch distance is at least the characteristic length
($R > \sigma/\sqrt{s_m}$),
and that
\textbf{(c)} local drift is not so strong that the equilibirum frequency is at least approximately attained
(using Wright's effective neighborhood size, $s_m \gg 1/(\rho \sigma^d)$).
The last requirement is necessary because if the $B$ allele fixes in large neighborhoods where it is deleterious,
we cannot approximate its dynamics via a branching process.
We also neglect the time for migration-selection equilibrium to be reached,
which has the effect of over-estimating $\migrate$.
As discussed above, we also assume that migration to a new patch takes place over a number of generations;
if there are sufficient rare, long-distance migration events
that would move between patches in a single hop,
this would require a different analysis.

@@

\begin{figure}[ht!]
  \begin{center}
      % made in analyze-sims.R
      \includegraphics{times-predicted-observed}
  \end{center}
  \caption{
      \textbf{(left)} 
      Time until adaptation by mutation in an isolated patch for different parameter combinations,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:mutrate} (horizontal axis).
      \textbf{(right)} 
      Time until adaptation by migration from an already occupied patch to a new patch,
      as observed in simulations (vertical axis),
      and as predicted from equation~\eqref{eqn:migrate}, with $C=5$ (horizontal axis).
      For each,
      parameter combinations and numbers of simulations are in supplemental tables \ref{stab:mutation_params} and \ref{stab:migration_params};
      shown are the median (points) and interquartile range (lines)
      of the time until the patch had at least 100 adapted individuals,
      for each unique parameter combination.
  }   \label{fig:sim_times}
\end{figure}

@@

The other key issue in thinking about the population genetic signal of adaptation in spatial patches 
is the process by which haplotypes are whittled down within patches, 
which plays into a number of aspects of haplotype sharing. 
The initial haplotype that sweeps within a patch 
will be dispersed over time by recombination.
Likewise, the haplotype that is shared between patches coadapted by migration
(equation~\ref{eqn:haplotype_length}) will also break down. 
However, a long time after the initial sweep,
we may still expect to find individuals within the patch sharing longer haplotypes around the selected locus
than with individuals elsewhere,
since selection against migrants 
decreases mean coalescence times within the patch near the selected locus.  %\citep{bartonpaper,charlesworths}.

@@

First we briefly describe the simulations we used for illustration and validation
(the \texttt{R} code used is provided as a supplement).
We simulated forward-time dynamics of the number of alleles of each type in a square grid (either one- or two-dimensional) of demes with fixed size $N$.
Each generation, each individual independently chose to reproduce or not with a probability depending on her type and location in the grid;
locally beneficial alleles were more likely to reproduce.
Each then produced a random, Poisson distributed number of offspring that each, independently, 
either remained in the same location with probabilty $1-m$ or
migrated a random number of steps with probability $m$
in a uniformly chosen cardinal direction;
for most simulations $m=0.2$ and
the probability of migrating $k$ steps was proportional to $2^{-k}$ for $1\le k \le 5$.
\linelabel{rr:sim_details}
Once migrants were distributed, each deme was uniformly resampled back down to $N$ individuals.
(Although we described the simulation in terms of individuals,
we kept track only of total numbers in an equivalent way.)

@@

Now consider the chance that $Z$ ever hits the new patch, which we call $h(x,S)$.
(In the main body, this was $h_A(x,y)$, here modified to parallel $u(x,S)$.)
If we make the approximation that $Z$ hits the new patch only if the trunk of $Z$ does,
and recall that $1-k_e(t)$ is the chance that $Z$ survives for $t$ generations,
then this is equal to
\begin{align}
    h(x,S) &\approx \int_0^\infty (1-k_e(t)) \P^x\{ \tau_S \in dt \} \\
           &= \int_0^\infty e^{s_m t} (1-k_e(t)) e^{- s_m t } \P^x\{ \tau_S \in dt \} \\
           &=\int_0^\infty e^{s_m t} (1-k_e(t)) \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
           &\approx \frac{1}{\E[K]} \int_0^\infty \P^x\{ \tau_S \in dt \, \& \, t < \tau_\dagger \} \\
    &= \frac{1}{\E[K]} \P^x[ \tau_S < \tau_\dagger ] .
\end{align}
Therefore, we have that
\begin{align} \label{eqn:boils_downto}
    h(x,S) \E[K] 
    \approx \P^x[ \tau_S < \tau_\dagger ] 
    = u(x,S) / g(A) .
\end{align}
